---
title: "BLANK"
author: "Jacob Gutierrez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
    theme: yeti
subtitle: "BLANK"
---

```{r knitr setup chunk, echo = F}
## Knitr setupchunk
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE, rows.print=20, eval = TRUE)
```


# Overview {.tabset .tabset-fade .tabset-pills}

This script is expected to run on {BLANK} See sessionInfo for more info.


## Purpose   


> **Assess BLANK** 

## Motivation

## Research Questions


# Analysis 

### Libraries


```{r libraries}
require(data.table)
require(Seurat)
#require(Azimuth)
require(annotables)
require(dplyr)
library(ggpubr)
require(stringr)
require(ggplot2)
library(BPCells)
```


```{r}
std_chrs <- c(c(1:22) , c("X","Y","MT"))
clean_grch38 <- annotables::grch38 %>% filter(chr %in% std_chrs)

## Only take oldest entrez id
clean_grch38 <- clean_grch38 %>% group_by(symbol) %>% arrange(desc(entrez)) %>% slice_head(n=1)
```


# Hu2020_study1

```{r,eval=F}
hu2020_mat <- fread("../data/public_aligned_data/study1_Hu2020/GSE139079_counts_benign.txt.gz") #%>% as.sparse()

#hu2020_mat[1:5,1:5] #%>% head()
mat_genes <-  hu2020_mat$V1
#rownames(hu2020_mat) <- mat_genes

hu2020_mat <- hu2020_mat %>% select(-V1) %>% as.matrix()%>% as.sparse()
rownames(hu2020_mat) <- mat_genes

colnames(hu2020_mat) <- colnames(hu2020_mat) %>% paste0("Hu#",.)


idx <- which(mat_genes %in% annotables::grch38$symbol)

use_ensmble <-clean_grch38[which(clean_grch38$symbol %in% mat_genes) , ]



nstart <- hu2020_mat %>% nrow()
hu2020_mat <- hu2020_mat[use_ensmble$symbol, ]
nfinal <-  hu2020_mat %>% nrow()
#rownames(hu2020_mat) <- use_ensmble$ensgene
sprintf("%i start > %i (%0.1f%%) final ensgene mapped", nstart, nfinal, (nfinal/nstart) * 100)

#use_ensmble %>% group_by(symbol) %>% count() %>% arrange(desc(n))
#Azimuth::ConvertGeneNames(mat_genes)

## Writing BPCells output 
write_matrix_dir(hu2020_mat, dir = "../data/bpcells_flatfiles/hu2020", overwrite = T)
```


```{r, eval = F}
hu2020_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/hu2020")
```




# Dihn2021_study2

```{r, eval = F}
all_h5s <- list.files("../data/public_aligned_data/study2_Dihn2021/", pattern = ".h5", full.names = T)
names(all_h5s) <- all_h5s %>% basename() %>% gsub("GSM[[:digit:]]+_","",.)
options(Seurat.object.assay.version = "v5")
all_mats <- all_h5s %>% lapply(Read10X_h5)


## Simplifying informative names 
names(all_mats) <- names(all_mats) %>% gsub("FT-*","",.) %>% gsub(".h5","",.) %>% paste0("Dihn","#",.)
all_mats <- all_mats %>% seq_along() %>% lapply(function(i){ 
  mat <- all_mats[[i]]
  colnames(mat) <- paste0(names(all_mats)[i],"#",colnames(all_mats[[i]]))
  return(mat)
  })

dihn2021_obj <- CreateSeuratObject(counts = all_mats,project = "dihn2021")
dihn2021_obj[["RNA"]] <- JoinLayers(dihn2021_obj[["RNA"]]) ## use fancy seurat v5 merge for cheap!
dihn2021_mat <- as(dihn2021_obj[["RNA"]]$counts, Class = "dgCMatrix") ## now spit out the sparse mat



## Subset usable Genes
mat_genes <- rownames(dihn2021_mat)
use_ensmble <-clean_grch38[which(clean_grch38$symbol %in% mat_genes) , ]


nstart <- dihn2021_mat %>% nrow()
dihn2021_mat <- dihn2021_mat[use_ensmble$symbol, ]
nfinal <-  dihn2021_mat %>% nrow()
#rownames(dihn2021_mat) <- use_ensmble$ensgene ## use only symbols? 

sprintf("%i start > %i (%0.1f%%) final ensgene mapped", nstart, nfinal, (nfinal/nstart) * 100)

## Writing BPCells output 
write_matrix_dir(dihn2021_mat, dir = "../data/bpcells_flatfiles/dihn2021",overwrite = T)
```



```{r}
dihn2021_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/dihn2021")
```


# Ulrich2022_study3


```{r, eval = F}
ulrich2022_obj <- readRDS("../data/public_aligned_data/study3_Ulrich2022/local.rds")



ulrich2022_mat <- as(ulrich2022_obj[["RNA"]]$counts, Class = "dgCMatrix") ## now spit out the sparse mat


## Forge new names

## Give a simple mapping to easily read
anatomical_map <- c("FM", "AMP", "IS", "FT", "FMAMP")
names(anatomical_map) <- c("Fimbria", "Ampulla", "Isthmus", "FT", "FimAmp")

donor_map <- colnames(ulrich2022_mat) %>% str_split("_") %>% lapply(function(x) x[[1]])  %>% unlist() %>% unique() %>% Reduce("rbind",.) %>% as.data.frame()
colnames(donor_map) <- "sample_id"
rownames(donor_map) <- NULL
donor_map$anatomical.full <- donor_map$sample_id %>% gsub("[[:digit:]]","",.) 
donor_map$donor_id <- donor_map$sample_id  %>% gsub("[Fimbria|Ampulla|Isthmus|FT|FimAmp]+","D",.)
donor_map$anatomical <- anatomical_map[donor_map$anatomical.full]
donor_map <- donor_map %>% mutate(jg_sample= sprintf("%s_%s",donor_id,anatomical))

fwrite(donor_map,"../data/public_aligned_data/study3_Ulrich2022/jg_sample_metadata.csv")


## Rename and write out names in convienet dataframe
#colnames(ulrich2022_mat)  <- colnames(ulrich2022_mat)  %>% gsub("_","#",.) %>% paste0("Ulrich#",.)
cb_df <- data.frame(cb = colnames(ulrich2022_mat))  %>% tidyr::separate(col = "cb", into = c("sample_id","cb"), sep = "_")#%>% as.data.frame()

cb_df <- left_join(cb_df, donor_map %>% select(sample_id, jg_sample)) 
cb_df <- cb_df %>% mutate(study = "Ulrich", new_cb = sprintf("%s#%s#%s", study, jg_sample, cb))

#colnames(ulrich2022_mat)  <- colnames(ulrich2022_mat)  %>% gsub("_","#",.) %>% paste0("Ulrich#",.)
colnames(ulrich2022_mat)  <- cb_df$new_cb



name_map <- data.frame(jg_cells = colnames(ulrich2022_mat), orig_cells = Cells(ulrich2022_obj))
meta_data <- cbind(name_map,ulrich2022_obj@meta.data)
fwrite(meta_data,"../data/public_aligned_data/study3_Ulrich2022/jg_ulrich_metadata.csv")





mat_ensmble <- rownames(ulrich2022_mat)
use_ensmble <-clean_grch38[which(clean_grch38$ensgene %in% mat_ensmble) , ]



nstart <- ulrich2022_mat %>% nrow()
ulrich2022_mat <- ulrich2022_mat[use_ensmble$ensgene, ]
nfinal <-  ulrich2022_mat %>% nrow()
rownames(ulrich2022_mat) <- use_ensmble$symbol ## use only symbols? 


sprintf("%i start > %i (%0.1f%%) final ensgene mapped", nstart, nfinal, (nfinal/nstart) * 100)


## Writing BPCells output 
write_matrix_dir(ulrich2022_mat, dir = "../data/bpcells_flatfiles/ulrich2022",overwrite = F)
```


```{r}
ulrich2022_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/ulrich2022")
```


#Lengyel2022_study4 


```{r, eval = F}

Lengyel2022_obj <- readRDS("../data/public_aligned_data/study4_Lengyel2022/local.rds")

#Lengyel2022_obj
#Lengyel2022_obj@meta.data %>% head()
Lengyel2022_mat <- as(Lengyel2022_obj[["RNA"]]$counts, Class = "dgCMatrix") ## now spit out the sparse mat


map_mat <- data.frame(orig_cells = colnames(Lengyel2022_mat), cellbc = colnames(Lengyel2022_mat) %>% gsub("(-1)?_[[:digit:]]","",.), lib_int = colnames(Lengyel2022_mat) %>% gsub("[[:alpha:]]+(-1)?_","",.), donor_id = Lengyel2022_obj$donor_id)

map_mat$clean_donor <- map_mat$donor_id %>% gsub("_[[:digit:]]+","",.)

map_mat <- map_mat %>% mutate(jg_cells = sprintf("Lengyel#%s_%s#%s", clean_donor,lib_int,cellbc))
map_mat <- map_mat %>% select(jg_cells, orig_cells)
meta_data <- cbind(map_mat, Lengyel2022_obj@meta.data)
fwrite(meta_data,"../data/public_aligned_data/study4_Lengyel2022/jg_lengyel_metadata.csv")


## Record name mapping for convience
woof <- c("AMP","FM","IS")
names(woof) <- c("ampulla of uterine tube", "fimbria of uterine tube", "isthmus of fallopian tube")

meta_data$sample_id <- meta_data$jg_cells %>% strsplit("#") %>% lapply(function(x)x[[2]]) %>% unlist()
sample_meta <- meta_data %>% distinct(sample_id,tissue )
sample_meta$donor_id <- sample_meta$sample_id %>% gsub("_[[:digit:]]+","",.)
sample_meta$anatomical <- woof[sample_meta$tissue]
sample_meta <- sample_meta %>% mutate(new_sample_id = sprintf("%s_%s", donor_id, anatomical))

fwrite(sample_meta,"../data/public_aligned_data/study4_Lengyel2022/jg_sample_metadata.csv")


## Rename cell barcodes with new names
colnames(Lengyel2022_mat) <- map_mat$jg_cells

## extract gene names in ensmbl format and directly map to ucsc 
mat_ensmble <- rownames(Lengyel2022_mat)
use_ensmble <-clean_grch38[which(clean_grch38$ensgene %in% mat_ensmble) , ]



## subset & count missing genes.
nstart <- Lengyel2022_mat %>% nrow()
Lengyel2022_mat <- Lengyel2022_mat[use_ensmble$ensgene, ]
nfinal <-  Lengyel2022_mat %>% nrow()
rownames(Lengyel2022_mat) <- use_ensmble$symbol ## use only symbols? 


sprintf("%i start > %i (%0.1f%%) final ensgene mapped", nstart, nfinal, (nfinal/nstart) * 100)

## Writing BPCells output 
write_matrix_dir(Lengyel2022_mat, dir = "../data/bpcells_flatfiles/lengyel2022",overwrite = F)
```


```{r}
lengyel2022_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/lengyel2022")
```


# Gutierrez et al 2023_study5

This study 


```{r, eval = F}
all_h5s <- list.files("../data/study_5_h5s/", pattern = ".h5", full.names = T)
names(all_h5s) <- all_h5s %>% basename() %>% gsub("_filtered_feature_bc_matrix.h5","",.)
options(Seurat.object.assay.version = "v5")
all_mats <- all_h5s %>% lapply(Read10X_h5)
the_s <- names(all_mats)


## Renaming donors using map and easy to read scheme
df <- data.frame(full_name = the_s)
df$patient_id <- df$full_name %>% gsub("-[[:alpha:]]+","",.)
df$anatomical <-  df$full_name %>% gsub("SHS-[[:digit:]]+-[[:digit:]]+-","",.)

mapping <- df %>% distinct(patient_id) %>% mutate(i =1 , rank = cumsum(i), donor = paste0("D",rank)) %>% select(patient_id, donor)
#donor_map <- mapping$donor
#names(donor_map) <- mapping$patient_id

df <- left_join(df,mapping)
df <- df %>% mutate(anatomical = if_else(anatomical == "Fimbria", "FM",anatomical), jg_donor = sprintf("%s_%s",donor,anatomical))

fwrite(df, "../data/study_5_h5s/jg_donor_map.csv")

## Renaming the thing to make pretty cell names
names(all_mats) <- df$jg_donor

all_mats <- all_mats %>% seq_along() %>% lapply(function(i) { 
  df <- all_mats[[i]]
  colnames(df) <- colnames(df) %>% paste0("Gutierrez#",names(all_mats)[i],"#",.)
  return(df)}) 



thisstudy_obj <- CreateSeuratObject(counts = all_mats,project = "gutierrez2023")
thisstudy_obj[["RNA"]] <- JoinLayers(thisstudy_obj[["RNA"]]) ## use fancy seurat v5 merge for cheap!
thisstudy_mat <- as(thisstudy_obj[["RNA"]]$counts, Class = "dgCMatrix") ## now spit out the sparse mat


mat_genes <- rownames(thisstudy_mat)
use_ensmble <-clean_grch38[which(clean_grch38$symbol %in% mat_genes) , ]



nstart <- thisstudy_mat %>% nrow()
thisstudy_mat <- thisstudy_mat[use_ensmble$symbol, ]
nfinal <-  thisstudy_mat %>% nrow()
#rownames(dihn2021_mat) <- use_ensmble$ensgene ## use only symbols? 

sprintf("%i start > %i (%0.1f%%) final ensgene mapped", nstart, nfinal, (nfinal/nstart) * 100)

## Writing BPCells output 
write_matrix_dir(thisstudy_mat, dir = "../data/bpcells_flatfiles/gutierrez2023",overwrite = F)
```


```{r}
gutierrez2023_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/gutierrez2023")
```




# Read in all BPCells matrices and naively forge 


```{r}
#ord <- c("Hu2020_study1","Dihn2021_study2","Ulrich2022_study3","Lengyel2022_study4","Gutierrez2023_study5")
#in_mats <-  list(hu2020_mat,dihn2021_mat,ulrich2022_mat,lengyel2022_mat,gutierrez2023_mat)

ord <- c("Dihn2021_study2","Ulrich2022_study3","Lengyel2022_study4","Gutierrez2023_study5")

in_mats <-  list(dihn2021_mat,ulrich2022_mat,lengyel2022_mat,gutierrez2023_mat)


 in_mats %>% lapply(rownames) %>% lapply(function(x) x[str_detect(x,"MT-")])



## Subset to genes shared across all matrices
#in_mats %>% lapply(function(x) x %>% nrow)
 
#in_mats %>% lapply(function(x) x %>% rownames() %>% head())
overlap_genes <- in_mats %>% lapply(rownames) %>% Reduce("intersect",.)

length(overlap_genes) ## 17693 gene symbols overlap 
#sum(rownames(in_mats[[1]]) %in% rownames(in_mats[[2]]) )

in_mats <- in_mats %>% lapply(function(x) x[which(rownames(x) %in% overlap_genes) , ])

## LOok at it
in_mats %>% lapply(function(x) x %>% rownames() %>% head())
in_mats %>% lapply(function(x) x %>% dim)

#in_mats %>% lapply(nrow)
```


## Doublet Detection


```{r, eval = F}
#dihn2021_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/dihn2021")
dihn_obj <- CreateSeuratObject(dihn2021_mat)
dihn_obj$sample_id <- Cells(dihn_obj) %>% strsplit("#") %>% lapply(function(x) x[[2]]) %>% unlist()


#dihn_obj <- scDblFinder(dihn_obj, samples="sample_id", BPPARAM=MulticoreParam(8))
sce <- scDblFinder(as(GetAssayData(dihn_obj, layer="counts"), Class = "dgCMatrix"), samples=dihn_obj$sample_id)


dihn_obj$scDblFinder.class <- sce$scDblFinder.class
dihn_obj$scDblFinder.score <- sce$scDblFinder.score

out_meta <- dihn_obj@meta.data  %>% select(sample_id,scDblFinder.class,scDblFinder.score )%>% tibble::rownames_to_column(var = "jg_cells")
fwrite(out_meta, "../data/public_aligned_data/study2_Dihn2021/jg_scdblfinder_meta.csv")
```









```{r, eval = F}
#gutierrez2023_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/gutierrez2023")

jg_obj <- CreateSeuratObject(gutierrez2023_mat)
jg_obj$sample_id <- Cells(jg_obj) %>% strsplit("#") %>% lapply(function(x) x[[2]]) %>% unlist()


#dihn_obj <- scDblFinder(dihn_obj, samples="sample_id", BPPARAM=MulticoreParam(8))
sce <- scDblFinder(as(LayerData(jg_obj, layer="counts"), Class = "dgCMatrix"), samples=jg_obj$sample_id, BPPARAM=MulticoreParam(8))


jg_obj$scDblFinder.class <- sce$scDblFinder.class
jg_obj$scDblFinder.score <- sce$scDblFinder.score

out_meta <- jg_obj@meta.data  %>% select(sample_id,scDblFinder.class,scDblFinder.score )%>% tibble::rownames_to_column(var = "jg_cells")
fwrite(out_meta, "../data/study_5_h5s/jg_scdblfinder_meta.csv")
```




```{r}
dihn_dbl <- fread("../data/public_aligned_data/study2_Dihn2021/jg_scdblfinder_meta.csv")

#dihn_dbl %>% count(sample_id,scDblFinder.class) 
dihn_dbl %>% group_by(sample_id) %>% count(scDblFinder.class)  %>% mutate(prop= n/sum(n), lprop = sprintf("%i (%0.1f%%)",n,prop*100))  %>% filter(scDblFinder.class == "doublet")  %>% arrange(desc(prop))%>% ggbarplot(x = "sample_id", y = "prop", fill = "scDblFinder.class", label = .$lprop, position = position_dodge2(.9), orient = "horiz", lab.vjust = .5,lab.hjust = -.1) + scale_y_continuous(expand = expansion(mult= c(0,.25)))

```



```{r}
jg_dbl <- fread( "../data/study_5_h5s/jg_scdblfinder_meta.csv")

jg_dbl %>% group_by(sample_id) %>% count(scDblFinder.class)  %>% mutate(prop= n/sum(n), lprop = sprintf("%i (%0.1f%%)",n,prop*100))  %>% filter(scDblFinder.class == "doublet")%>% arrange(desc(prop)) %>% ggbarplot(x = "sample_id", y = "prop", fill = "scDblFinder.class", label = .$lprop, position = position_dodge2(.9), orient = "horiz", lab.vjust = .5,lab.hjust = -.1) + scale_y_continuous(expand = expansion(mult= c(0,.25)))
```


```{r}
dbl_df <- rbind(dihn_dbl,jg_dbl) %>% as.data.frame()

dbl_df <- dbl_df %>% select(-sample_id,-scDblFinder.score)
rownames(dbl_df) <- dbl_df$jg_cells

dbl_df <- dbl_df %>% select(-jg_cells)
```





## Naive Clustering



```{r, eval = F}
## Forge Object and shove in some metadata
comb.obj <- CreateSeuratObject(counts = in_mats)

meta_df <- seq_along(ord) %>% lapply(function(i) {
  data.frame(cb=colnames(in_mats[[i]]), study = ord[i]) 
}) %>% Reduce("rbind",.)

rownames(meta_df) <- meta_df$cb
meta_df <- meta_df %>% select(-cb)

comb.obj$study <- meta_df$study


# Naive Clustering no correction of atoms
obj <- comb.obj

## Doublets
obj <- AddMetaData(obj, metadata = dbl_df, col.name= "scDblFinder.class") 
obj$scDblFinder.class[is.na(obj$scDblFinder.class)] <- "singlet" ## Manually set all NA's to be singlet because the data is already filtered for singlets.... 

### ROUND 2 
DefaultAssay(obj) <- "RNA"
#obj <- obj %>% ScaleData()
obj <- obj %>% PercentageFeatureSet(pattern = "^MT-", col.name = "perc.mito")


assess_obj <- function(an_obj,the_col = "raw"){
  an_obj@meta.data %>% group_by(study) %>% count() %>% mutate(type = the_col)
}

out_cells <- obj %>% assess_obj()

obj <- obj %>% subset(subset = nFeature_RNA >= 500 )

two <- obj %>% assess_obj(the_col = ">=500RNAs")
out_cells <- rbind(out_cells, two)
  
# obj <- obj %>% subset(subset = nCount_RNA >= 1000 )
# 
# two <- obj %>% assess_obj(the_col = ">=1000counts")
# out_cells <- rbind(out_cells, two)
  
  
obj <- obj %>% subset(subset = perc.mito < 20)

two <- obj %>% assess_obj(the_col = "<20%mito")
out_cells <- rbind(out_cells, two)


obj <- obj %>% subset(subset = scDblFinder.class == "singlet")

two <- obj %>% assess_obj(the_col = "singlet")
out_cells <- rbind(out_cells, two)
fwrite(out_cells, "../output/processing/cellcounts_processing.csv")

## ENd of ROUnd 2 

## Attempt to set usable names? 
obj <- JoinLayers(obj)

obj[["RNA"]] <- split(obj[["RNA"]], f = obj$study)


obj <- NormalizeData(obj)
obj <- FindVariableFeatures(obj)
## Sample 10k cells per study to inform the integration
obj <- SketchData(
  object = obj,
  ncells = 10000,
  method = "LeverageScore",
  sketched.assay = "sketch"
)
#obj


DefaultAssay(obj) <- "sketch"
obj <- FindVariableFeatures(obj)
obj <- ScaleData(obj)
obj <- RunPCA(obj)
obj <- FindNeighbors(obj, dims = 1:30)
obj <- FindClusters(obj, resolution = 1)
obj <- RunUMAP(obj, dims = 1:30, return.model = T) ## I forgot why this is broken... but you need to bash `mkdir /tmp/RtmpIwVFPj/{filec0a5f633489}` based on the console output.

## RND 1
#saveRDS(obj, "../output/objects/combined_ft_atlas_raw.rds")

## RND2
#saveRDS(obj, "../output/objects/combined_ft_atlas_raw_rnd2.rds")


## RND3 BC SEURAT WAS UPDATED AAA
saveRDS(obj, "../output/objects/combined_ft_atlas_raw_rnd3.rds")

## Updated filtering for 1000 umis but rolling back bc it didn't impact analytics workflow 
#saveRDS(obj, "../output/objects/combined_ft_atlas_raw_rnd4.rds")
```


```{r, eval = F}
obj <- readRDS("../output/objects/combined_ft_atlas_raw_rnd2.rds")

```



## FFPE + Fresh Naive Clustering


# Gutierrez FFPE

This study 


```{r, eval = F}


ffpe.obj <- readRDS("../../FT_FFPE_snRNA/output/objects/raw_celltype_obj.rds")

thisstudy_mat <- as(ffpe.obj[["RNA"]]$counts, Class = "dgCMatrix") ## now spit out the sparse mat

colnames(thisstudy_mat) <- sprintf("Gutierrez_FFPE#%s",colnames(thisstudy_mat))

#thisstudy_obj <- CreateSeuratObject(counts = thisstudy_mat,project = "Gutierrez_FFPE")
#thisstudy_obj[["RNA"]] <- JoinLayers(thisstudy_obj[["RNA"]]) ## use fancy seurat v5 merge for cheap!
#thisstudy_mat <- as(thisstudy_obj[["RNA"]]$counts, Class = "dgCMatrix") ## now spit out the sparse mat


mat_genes <- rownames(thisstudy_mat)
use_ensmble <-clean_grch38[which(clean_grch38$symbol %in% mat_genes) , ]



nstart <- thisstudy_mat %>% nrow()
thisstudy_mat <- thisstudy_mat[use_ensmble$symbol, ]
nfinal <-  thisstudy_mat %>% nrow()
#rownames(dihn2021_mat) <- use_ensmble$ensgene ## use only symbols? 

sprintf("%i start > %i (%0.1f%%) final ensgene mapped", nstart, nfinal, (nfinal/nstart) * 100)

## Writing BPCells output 
write_matrix_dir(thisstudy_mat, dir = "../data/bpcells_flatfiles/ffpe_gutierrez2023",overwrite = F)
```


```{r}
ffpe_mat <- BPCells::open_matrix_dir("../data/bpcells_flatfiles/ffpe_gutierrez2023")
```





# Read in all BPCells matrices and naively forge 


```{r}
#ord <- c("Hu2020_study1","Dihn2021_study2","Ulrich2022_study3","Lengyel2022_study4","Gutierrez2023_study5")
#in_mats <-  list(hu2020_mat,dihn2021_mat,ulrich2022_mat,lengyel2022_mat,gutierrez2023_mat)

ord <- c("Dihn2021_study2","Ulrich2022_study3","Lengyel2022_study4","Gutierrez2023_study5","Gutierrez2023_ffpe")

in_mats <-  list(dihn2021_mat,ulrich2022_mat,lengyel2022_mat,gutierrez2023_mat,ffpe_mat)


 in_mats %>% lapply(rownames) %>% lapply(function(x) x[str_detect(x,"MT-")])



## Subset to genes shared across all matrices
#in_mats %>% lapply(function(x) x %>% nrow)
 
#in_mats %>% lapply(function(x) x %>% rownames() %>% head())
overlap_genes <- in_mats %>% lapply(rownames) %>% Reduce("intersect",.)

length(overlap_genes) ## 17693 gene symbols overlap 
#sum(rownames(in_mats[[1]]) %in% rownames(in_mats[[2]]) )

in_mats <- in_mats %>% lapply(function(x) x[which(rownames(x) %in% overlap_genes) , ])

## LOok at it
in_mats %>% lapply(function(x) x %>% rownames() %>% head())
in_mats %>% lapply(function(x) x %>% dim)

#in_mats %>% lapply(nrow)
```



```{r, eval = F}
## Forge Object and shove in some metadata
comb.obj <- CreateSeuratObject(counts = in_mats)

meta_df <- seq_along(ord) %>% lapply(function(i) {
  data.frame(cb=colnames(in_mats[[i]]), study = ord[i]) 
}) %>% Reduce("rbind",.)

rownames(meta_df) <- meta_df$cb
meta_df <- meta_df %>% select(-cb)

comb.obj$study <- meta_df$study


# Naive Clustering no correction of atoms
obj <- comb.obj

## Doublets
obj <- AddMetaData(obj, metadata = dbl_df, col.name= "scDblFinder.class") 
obj$scDblFinder.class[is.na(obj$scDblFinder.class)] <- "singlet" ## Manually set all NA's to be singlet because the data is already filtered for singlets.... 

### ROUND 2 
DefaultAssay(obj) <- "RNA"
#obj <- obj %>% ScaleData()
obj <- obj %>% PercentageFeatureSet(pattern = "^MT-", col.name = "perc.mito")


assess_obj <- function(an_obj,the_col = "raw"){
  an_obj@meta.data %>% group_by(study) %>% count() %>% mutate(type = the_col)
}

out_cells <- obj %>% assess_obj()

obj <- obj %>% subset(subset = nFeature_RNA >= 500 )

two <- obj %>% assess_obj(the_col = ">=500RNAs")
out_cells <- rbind(out_cells, two)
  
# obj <- obj %>% subset(subset = nCount_RNA >= 1000 )
# 
# two <- obj %>% assess_obj(the_col = ">=1000counts")
# out_cells <- rbind(out_cells, two)
  
  
obj <- obj %>% subset(subset = perc.mito < 20)

two <- obj %>% assess_obj(the_col = "<20%mito")
out_cells <- rbind(out_cells, two)


obj <- obj %>% subset(subset = scDblFinder.class == "singlet")

two <- obj %>% assess_obj(the_col = "singlet")
out_cells <- rbind(out_cells, two)
fwrite(out_cells, "../output/processing/cellcounts_processing.csv")

## ENd of ROUnd 2 

## Attempt to set usable names? 
obj <- JoinLayers(obj)

obj[["RNA"]] <- split(obj[["RNA"]], f = obj$study)


obj <- NormalizeData(obj)
obj <- FindVariableFeatures(obj)
## Sample 10k cells per study to inform the integration
obj <- SketchData(
  object = obj,
  ncells = 10000,
  method = "LeverageScore",
  sketched.assay = "sketch"
)
#obj


DefaultAssay(obj) <- "sketch"
obj <- FindVariableFeatures(obj)
obj <- ScaleData(obj)
obj <- RunPCA(obj)
obj <- FindNeighbors(obj, dims = 1:30)
obj <- FindClusters(obj, resolution = 1)
obj <- RunUMAP(obj, dims = 1:30, return.model = T) ## I forgot why this is broken... but you need to bash `mkdir /tmp/RtmpIwVFPj/{filec0a5f633489}` based on the console output.

## RND 1
saveRDS(obj, "../output/objects/ffpe_combined_ft_atlas_raw.rds")

```



# SessionInfo
<details>
  <summary>Click for SessionInfo</summary>
```{r R info}
sessionInfo()
```
</details> 

