---
title: "DownloadFT_scRNA"
author: "Jacob Gutierrez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
    theme: yeti
subtitle: "BLANK"
---

```{r knitr setup chunk, echo = F}
## Knitr setupchunk
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE, rows.print=20, eval = TRUE)
```


# Overview {.tabset .tabset-fade .tabset-pills}

This script is expected to run on {BLANK} See sessionInfo for more info.


## Purpose   


> **Assess BLANK** 

## Motivation

## Research Questions


# Analysis 

### Libraries


```{r libraries}
#library(rjson)
library(tidyverse)
library(GEOquery)
library(openxlsx)
```



How I get the data
http://seandavi.github.io/GEOquery/articles/GEOquery.html





# Hu2020_study1

Cancer: GSE132149
Benign: GSE139079


https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE139079


```{r}

wow_name <- "study1_Hu2020"
dir.create(wow_name,showWarnings = F)
wow_s <- "GSE139079"
gds <- getGEO(wow_s)
gsm <- gds[[paste0(wow_s,"_series_matrix.txt.gz")]]
cnt_url <- gsm@experimentData@other$supplementary_file
download.file(url = cnt_url,destfile = sprintf("%s/%s",wow_name,basename(cnt_url)),method="wget")
```


# Dihn2021_study2

https://pubmed.ncbi.nlm.nih.gov/33852846/

Dinh HQ, Lin X, Abbasi F, Nameki R, Haro M, Olingy CE, Chang H, Hernandez L, Gayther SA, Wright KN, Aspuria PJ, Karlan BY, Corona RI, Li A, Rimel BJ, Siedhoff MT, Medeiros F, Lawrenson K. Single-cell transcriptomics identifies gene expression networks driving differentiation and tumorigenesis in the human fallopian tube. Cell Rep. 2021 Apr 13;35(2):108978. doi: 10.1016/j.celrep.2021.108978. PMID: 33852846; PMCID: PMC10108902.

scRNA: GSE151214

> 10x 3' and DROP-seq 


```{r}
wow_name <- "study2_Dihn2021"
dir.create(wow_name,showWarnings = F)
wow_s <- "GSE151214"
gds <- getGEO(wow_s)
gsm <- gds[[paste0(wow_s,"_series_matrix.txt.gz")]] 
gsm_acc_df <- gsm %>% phenoData() %>% .@data
cnt_url <- gsm@experimentData@other$supplementary_file ## Tar with the .h5 files

#gsm_acc_df <- gds$GSE151214_series_matrix.txt.gz %>% phenoData() %>% .@data


want_acc_df <- gsm_acc_df %>% dplyr::select(title,geo_accession,patient_id = `individual:ch1`, source_name_ch1,tissue =  `tissue:ch1`,h5_path = `supplementary_file_1`)

```


Now download

```{r}

#meta_path <- paste0(out_dir,"dinh_et_al_2021_metadata.csv")
meta_path <- sprintf("%s/%s_metadata.csv",wow_name,wow_name)
write_csv(want_acc_df,meta_path)


#wget_str <- sprintf("wget %s",want_acc_df$h5_path)
#wget_path  <- paste0(out_dir,"scrna_h5_wget.sh")
#writeLines(wget_str,wget_path)

cnt_url <- gsm@experimentData@other$supplementary_file
download.file(url = cnt_url,destfile = sprintf("%s/%s",wow_name,basename(cnt_url)),method="wget")
untar(tarfile =  sprintf("%s/%s",wow_name,basename(cnt_url)),exdir = wow_name)

```


# Ulrich2022_study3


https://pubmed.ncbi.nlm.nih.gov/35320732/

Ulrich ND, Shen YC, Ma Q, Yang K, Hannum DF, Jones A, Machlin J, Randolph JF Jr, Smith YR, Schon SB, Shikanov A, Marsh EE, Lieberman R, Gurczynski SJ, Moore BB, Li JZ, Hammoud S. Cellular heterogeneity of human fallopian tubes in normal and hydrosalpinx disease states identified using scRNA-seq. Dev Cell. 2022 Apr 11;57(7):914-929.e7. doi: 10.1016/j.devcel.2022.02.017. Epub 2022 Mar 22. PMID: 35320732; PMCID: PMC9007916.

scRNA: GSE178101

cellxgene: https://cellxgene.cziscience.com/collections/fc77d2ae-247d-44d7-aa24-3f4859254c2c

```{r}
wow_name <- "study3_Ulrich2022"
dir.create(wow_name,showWarnings = F)
wow_s <- "GSE178101"
gds <- getGEO(wow_s)
gsm <- gds[[paste0(wow_s,"_series_matrix.txt.gz")]] 
gsm_acc_df <- gsm %>% phenoData() %>% .@data
cnt_url <- gsm@experimentData@other$supplementary_file ## Tar with the .h5 files
want_acc_df <- gsm_acc_df %>% dplyr::select(title,geo_accession,description,age=`age:ch1`,bmi=`bmi:ch1`,health=`health:ch1`)


#gsm_acc_df <- gds$GSE151214_series_matrix.txt.gz %>% phenoData() %>% .@data
```


```{r, eval = F}
## THE HEALTHY ONLY NOPE!
#curl -o local.rds "https://corpora-data-prod.s3.amazonaws.com/01d2bc41-c26d-4020-ba4b-f3c29a05dbe8/local.rds?AWSAccessKeyId=ASIATLYQ5N5X3WJDM5VJ&Signature=QOsOyTvArM9pNKmFb5HPKPo5vb0%3D&x-amz-security-token=IQoJb3JpZ2luX2VjEGUaCXVzLXdlc3QtMiJIMEYCIQCn%2BOmaIpjVwI5%2B6pjiZYB6H39Ns8vacy8pu9HIbJZnaAIhAOxj7EoPWDQCo0kU1LNEmX3kUma0zBIlpCsW0tdj4FPGKusDCH4QARoMMjMxNDI2ODQ2NTc1IgxrjuvJNcLp00AjRCoqyAPDkdxWNx6Zw6Cjppox9pM5%2FTL6bfyUqg%2BEreXYJ4Iy%2FIIhpQuE%2FvNjlgV9o3v7FlCyatl%2FCRO3n%2BA1HOoJjXV1NxYbd9W3%2BF1Jj1G6B2xv%2FCeMikrRcu7nvD0CLLPcate6nkHh8w85IVoJ0SMi4DHgrIbhuud92HeMDwb28RXdyehxXw1D6msFdAo6V1%2FhkHW1gVDdvgeeNKZTJxfiCufEdWukuDh67n8kV8lq6lxzmaVRc2LrY2fmRXnXrk67r%2BG%2Bhti%2F5bO7mokwjWu6I3nubo48iRJ9jLh6kovPx4qzUsR2CK82t1mfFHlVhvUtndenQWEIgW%2BxtL4dQr%2BoBdBN46KN%2BvGbkw6flGpp8eDkYN5XhkD4iUqiEBDeIGL6taF65k1k%2FCp%2FKIsiEcopBOKBnxHltzRz%2B8O2XszNm5hzyureFHfZfT7RIxdkZnEOY%2BBbe66UNcwZVgTncSd5DXVtBq6rRQjv1I2zIOnrphwyAwTuaaEf2lPbrKp0FIeKdd8BHnOglwXVeZ%2BzCQMqepLJ5zqb0SoeGYUD7%2FdWa0kh8j1CcqjZYCJX3bWd5UNgZc0ASFssTxb1vtTdHOZGVWzQngDlFdDyzwwwlOTqogY6pAGhz0KKcHwo4IvDjfoDEP7uPXpjot9iu%2B5vNBTY0m%2BWhMcOnrbNQQ9tGE83W21J0Yu7bebcElRPI1fy5iPgSVMILOfJAsjcPycbGIhAh2Hq3c37kVlwwcXt0Pi9Jfz7cGO%2FehZEsJOasv3zPuNT3cY5oK7dys1bdNsHd4o4IPLhOtnRVRrLbOigHqQMuABzBHsw3v2NSrX2wDeB1e%2F9G23mj6mj%2BQ%3D%3D&Expires=1684287774"

# ALL DATA YES?
#curl -o local.rds "https://corpora-data-prod.s3.amazonaws.com/3549f484-9650-4dd4-b0b4-4f50ab6f7e06/local.rds?AWSAccessKeyId=ASIATLYQ5N5XV2SFORFD&Signature=VNn8eaw%2BidMOMoKcJnN6qw6iO4M%3D&x-amz-security-token=IQoJb3JpZ2luX2VjEG4aCXVzLXdlc3QtMiJHMEUCIFp2LVFlfbTfY9PJbXRtTkFGHNVnQFRqGhSnZcMUnHRMAiEAnPN1gtx06TOT6GJ5xUL55MXhgqlv5e5IduRzPzFPouUq9AMIh%2F%2F%2F%2F%2F%2F%2F%2F%2F%2F%2FARABGgwyMzE0MjY4NDY1NzUiDJZY2kpEZvZiJcU1yirIA1slDEHPJD8x8tPQbLGLGU0YTK1t9JPJ8vjU6Ly%2Fo3MDFwfgkXdPSmhQ3BdnzR9EDb1j2ZyWpMomRfeielYaIiDMiVFfRKYbSy%2FM%2F4DdTUOL4qXfzncynkCpjWtaoeMS8DjIvnO5BkloQuQI3cPmBhA7AgObiZ45JsxAGiBqZazgBhKWEbqaSo29r1y9dEErRO3YGgwJSVBlrX1yGutncft6BldeZrRAp4eowBBhUOlAWV1cgcg0TCLP9jnnLmHd6IhQ08zskcpc84nuRdrxMXOfYN%2Bk%2FXPL0RL0GOr%2B8RQnumwJ9fv0zBoXcTA5%2Bh7%2BfLQ9CIdqMeUenO46a4cYUNFKm2%2FG9ErW36fzW5AITQpU4Sv%2B0ZaaODTgxvWKXEQdEjaAsR6a9HV6uRGkBndSAebd88F2KIlpiSoHTfw%2FvaoptJEm9f3C3QGWu9J6qDs5nz7iXV1xOo6vyL5g6Xk6LdtMuRZMVwcX1%2BDkSzN92HUVdjPI%2BYVbnJsNErnX06uu7iAd4DNvlBViNwdENaOKmOSgdYkUert20Vkj5%2B1D7%2Bs3xIiY997jEaaMkchO3YjVjkr3chtWfjSRNcJ4c0wZCKAJpbCGJvLI7TDK3%2ByiBjqlATnLNSE8ddktUB%2F1OLgnvyL3e1hqQt1HFpwrl8EsCyh6VbeyIfwMNWY4Y%2B6Ewp1uvD3wP9lEPjRgP0tMd5%2B1c2ASZldBOxFpAdS2hfBKa7AJbXoG0bSn1qlOoxlXu378y4MWRHGFpuxwWGWeY%2B4UyuSjmsc%2B0jrLHVJ6B7Qf6QwfW2jw%2F79LfLngAXV6LOneLg3S%2F3v73nFqm3UuQVSyQsvVAP4lDQ%3D%3D&Expires=1684307563"
the_path <- "https://corpora-data-prod.s3.amazonaws.com/3549f484-9650-4dd4-b0b4-4f50ab6f7e06/local.rds?AWSAccessKeyId=ASIATLYQ5N5XV2SFORFD&Signature=VNn8eaw%2BidMOMoKcJnN6qw6iO4M%3D&x-amz-security-token=IQoJb3JpZ2luX2VjEG4aCXVzLXdlc3QtMiJHMEUCIFp2LVFlfbTfY9PJbXRtTkFGHNVnQFRqGhSnZcMUnHRMAiEAnPN1gtx06TOT6GJ5xUL55MXhgqlv5e5IduRzPzFPouUq9AMIh%2F%2F%2F%2F%2F%2F%2F%2F%2F%2F%2FARABGgwyMzE0MjY4NDY1NzUiDJZY2kpEZvZiJcU1yirIA1slDEHPJD8x8tPQbLGLGU0YTK1t9JPJ8vjU6Ly%2Fo3MDFwfgkXdPSmhQ3BdnzR9EDb1j2ZyWpMomRfeielYaIiDMiVFfRKYbSy%2FM%2F4DdTUOL4qXfzncynkCpjWtaoeMS8DjIvnO5BkloQuQI3cPmBhA7AgObiZ45JsxAGiBqZazgBhKWEbqaSo29r1y9dEErRO3YGgwJSVBlrX1yGutncft6BldeZrRAp4eowBBhUOlAWV1cgcg0TCLP9jnnLmHd6IhQ08zskcpc84nuRdrxMXOfYN%2Bk%2FXPL0RL0GOr%2B8RQnumwJ9fv0zBoXcTA5%2Bh7%2BfLQ9CIdqMeUenO46a4cYUNFKm2%2FG9ErW36fzW5AITQpU4Sv%2B0ZaaODTgxvWKXEQdEjaAsR6a9HV6uRGkBndSAebd88F2KIlpiSoHTfw%2FvaoptJEm9f3C3QGWu9J6qDs5nz7iXV1xOo6vyL5g6Xk6LdtMuRZMVwcX1%2BDkSzN92HUVdjPI%2BYVbnJsNErnX06uu7iAd4DNvlBViNwdENaOKmOSgdYkUert20Vkj5%2B1D7%2Bs3xIiY997jEaaMkchO3YjVjkr3chtWfjSRNcJ4c0wZCKAJpbCGJvLI7TDK3%2ByiBjqlATnLNSE8ddktUB%2F1OLgnvyL3e1hqQt1HFpwrl8EsCyh6VbeyIfwMNWY4Y%2B6Ewp1uvD3wP9lEPjRgP0tMd5%2B1c2ASZldBOxFpAdS2hfBKa7AJbXoG0bSn1qlOoxlXu378y4MWRHGFpuxwWGWeY%2B4UyuSjmsc%2B0jrLHVJ6B7Qf6QwfW2jw%2F79LfLngAXV6LOneLg3S%2F3v73nFqm3UuQVSyQsvVAP4lDQ%3D%3D&Expires=1684307563"

download.file(url = the_path,destfile = sprintf("%s/local.rds",wow_name),method="wget")




#meta_path <- paste0(out_dir,"dinh_et_al_2021_metadata.csv")
meta_path <- sprintf("%s/%s_metadata.csv",wow_name,wow_name)
write_csv(want_acc_df,meta_path)


```



#Lengyel2022_study4 

https://pubmed.ncbi.nlm.nih.gov/36543131/

Lengyel E, Li Y, Weigert M, Zhu L, Eckart H, Javellana M, Ackroyd S, Xiao J, Olalekan S, Glass D, Iyer S, Krishnan R, Bilecz AJ, Lastra R, Chen M, Basu A. A molecular atlas of the human postmenopausal fallopian tube and ovary from single-cell RNA and ATAC sequencing. Cell Rep. 2022 Dec 20;41(12):111838. doi: 10.1016/j.celrep.2022.111838. PMID: 36543131.

scRNA: https://cellxgene.cziscience.com/collections/d36ca85c-3e8b-444c-ba3e-a645040c6185


```{r}

wow_name <- "study4_Lengyel2022"
dir.create(wow_name,showWarnings = F)
#curl -o local.rds "https://corpora-data-prod.s3.amazonaws.com/a3d49cdd-18f1-4ed6-bd41-591356a9c1bc/local.rds?AWSAccessKeyId=ASIATLYQ5N5X2VDJQTR7&Signature=1QgcP1kDH5Xd4u2ezNZ7vtNzgkY%3D&x-amz-security-token=IQoJb3JpZ2luX2VjEG0aCXVzLXdlc3QtMiJHMEUCIQC1bU8PIuIBEDMejVBVWsBW3okFiFxz1IsAUDzSpIwiuwIgWw%2BxOYfcVCdWcAIDmBXIQehDk818zkahDGy4aJo%2FsXEq9AMIhv%2F%2F%2F%2F%2F%2F%2F%2F%2F%2FARABGgwyMzE0MjY4NDY1NzUiDGr7l797TmirTp7ClyrIA1vD2BJQIbux28pMBG613ccCNgvOAmydlkY5%2FOiF6SWWTyj1C1CV2euRhznvBsq8tD%2BcY4AkbpbOyIbS%2Fzbd1O4CEo%2FlMHTlPjBuazxiedfEGRHrC%2BA0jlD0WV%2BjVfYRQdXnNzh31TzXYZd%2BfT04cFlKJ7T3fDpKIC1ZWWeVH1CHUgzK3W9KGTYSo%2F7732Sb6UbNMr2qL8%2BQlJJE9B9QKQQ8NjVVADGh7o5Ds4m1SUCqGUgADY2I%2F6rG6Xwr1w%2ByuYrOJwkh9xmX78L5VJBem870ReLgNijkaqOVuKFJescFvLTFqchCvuLiVAcXwuCqRk87W2gbeXocTkSXS70bajW6XFbMiDHSSuAWWUthjAUHunHotGRgyJHW1yJZ2Zt8810GjU%2F3Y8Dl3y9%2FQYxjzJsl8CDCmIVYCd5PKgWlsmZgpCDX4cSjMyrmfTC7yJXaxmjq7JJsPhNX4EacZCbkywFt9o8fX7mqXDom6lXCz6liMIVU4%2BqVcGwhDEFHUXtA0rNMZTsWff5osi4vbK7S%2BMKEGBFZ%2BTW841PKdw0Lt7n6c0W4wK%2Bn6dR6r2nh%2F7bSyCOKywhVAGcdA8CEFW3jwVTBXIFrLN4msTDYvuyiBjqlASfST5dLfBJoEPUBXHFOY7Mct1uHtSFliBSR1Nl37ZWd2bDRrvh6HbHVYNmFppTNKIeD%2FZmNrGYUXoUsSl4mP2mE5YOsKd%2BoaWUwCJ5ZYEDhI9ghVsDuf4UvpHgjwXLDjZlrBvjYrWSWuIHAbbMVrIJJ4NUAObZWj7RFyYmlqqGO0LJjo6xPGyr8PckFEz9724a1PD2X5sn2A%2BgYw1qfvk7ICEvWTA%3D%3D&Expires=1684307866"


the_path <- "https://corpora-data-prod.s3.amazonaws.com/a3d49cdd-18f1-4ed6-bd41-591356a9c1bc/local.rds?AWSAccessKeyId=ASIATLYQ5N5X2VDJQTR7&Signature=1QgcP1kDH5Xd4u2ezNZ7vtNzgkY%3D&x-amz-security-token=IQoJb3JpZ2luX2VjEG0aCXVzLXdlc3QtMiJHMEUCIQC1bU8PIuIBEDMejVBVWsBW3okFiFxz1IsAUDzSpIwiuwIgWw%2BxOYfcVCdWcAIDmBXIQehDk818zkahDGy4aJo%2FsXEq9AMIhv%2F%2F%2F%2F%2F%2F%2F%2F%2F%2FARABGgwyMzE0MjY4NDY1NzUiDGr7l797TmirTp7ClyrIA1vD2BJQIbux28pMBG613ccCNgvOAmydlkY5%2FOiF6SWWTyj1C1CV2euRhznvBsq8tD%2BcY4AkbpbOyIbS%2Fzbd1O4CEo%2FlMHTlPjBuazxiedfEGRHrC%2BA0jlD0WV%2BjVfYRQdXnNzh31TzXYZd%2BfT04cFlKJ7T3fDpKIC1ZWWeVH1CHUgzK3W9KGTYSo%2F7732Sb6UbNMr2qL8%2BQlJJE9B9QKQQ8NjVVADGh7o5Ds4m1SUCqGUgADY2I%2F6rG6Xwr1w%2ByuYrOJwkh9xmX78L5VJBem870ReLgNijkaqOVuKFJescFvLTFqchCvuLiVAcXwuCqRk87W2gbeXocTkSXS70bajW6XFbMiDHSSuAWWUthjAUHunHotGRgyJHW1yJZ2Zt8810GjU%2F3Y8Dl3y9%2FQYxjzJsl8CDCmIVYCd5PKgWlsmZgpCDX4cSjMyrmfTC7yJXaxmjq7JJsPhNX4EacZCbkywFt9o8fX7mqXDom6lXCz6liMIVU4%2BqVcGwhDEFHUXtA0rNMZTsWff5osi4vbK7S%2BMKEGBFZ%2BTW841PKdw0Lt7n6c0W4wK%2Bn6dR6r2nh%2F7bSyCOKywhVAGcdA8CEFW3jwVTBXIFrLN4msTDYvuyiBjqlASfST5dLfBJoEPUBXHFOY7Mct1uHtSFliBSR1Nl37ZWd2bDRrvh6HbHVYNmFppTNKIeD%2FZmNrGYUXoUsSl4mP2mE5YOsKd%2BoaWUwCJ5ZYEDhI9ghVsDuf4UvpHgjwXLDjZlrBvjYrWSWuIHAbbMVrIJJ4NUAObZWj7RFyYmlqqGO0LJjo6xPGyr8PckFEz9724a1PD2X5sn2A%2BgYw1qfvk7ICEvWTA%3D%3D&Expires=1684307866"

download.file(url = the_path,destfile = sprintf("%s/local.rds",wow_name),method="wget")





```




# SessionInfo
<details>
  <summary>Click for SessionInfo</summary>
```{r R info}
sessionInfo()
```
</details> 

